Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-09-10 main.tex
:pencil: Update manuscript

			denotes the typical received signal energy over $N$ symbol blocks when the tag is at state $m$, which can be estimated at the user. For the ease of exposition, we sort $\{\sigma_m^2\}$ in a descending order by bijective mapping $f \colon m \mapsto i$\footnote{Note that the mapping is not unique and the detection fails when two different constellation points yield the same energy level.} and let $\sigma_i^2 \ge \sigma_j^2$ for $i > j$, $i,j \in \mathcal{M}$. Denote $\mathcal{H}_i$ as the hypothesis that the tag transmits $c_i$. Following \cite{Qian2019}, the optimal Maximum-Likelihood (ML) detector boils down to the energy detector, and the likelihood ratio between hypotheses $\mathcal{H}_i$ and $\mathcal{H}_j$ is

@@

				\Lambda_{i, j}(\boldsymbol{y}) \underset{H_j}{\overset{H_i}{\gtrless}} 1 \iff z \underset{H_j}{\overset{H_i}{\gtrless}} T_{i, j},

@@

			We then investigate the achievable rate of the backscatter link. Denote the tag input probability mass function as $\boldsymbol{r} = [r_1,\ldots,r_M]^T \in \mathbb{R}_+^{M \times 1}$ where $r_m$ is the probability to transmit symbol $c_m$ [strictly positive?]. For a given input distribution $\boldsymbol{r}$ and forward transition matrix $\boldsymbol{P}$, the achievable backscatter rate can be written as
			\begin{equation}
				R_c = \max_{\boldsymbol{Q} \ge 0} \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i} \ [\si{bpcu}],
				\label{eq:secondary_rate}
			\end{equation}
			where $\boldsymbol{Q} \in \mathbb{R}_{+}^{M \times M}$ is the backward transition probability matrix \cite{Yeung2008}
			\begin{equation}
				\boldsymbol{Q} \triangleq
				\begin{bmatrix}
					q_{1, 1} & \ldots & q_{1, M} \\
					\vdots & \ddots & \vdots \\
					q_{M, 1} & \ldots & q_{M, M}
				\end{bmatrix},
			\end{equation}
			and $q_{j, i}$ denotes the transition probability from channel output state $j$ to input state $i$.
		\end{subsection}

@@

	\begin{section}{Tag Input Distribution}
		To achieve a flexible tradeoff between the backscatter and legacy links, we aims to optimize the tag input probability mass function, under $M$-PSK modulation, to maximize the weighted sum of the primary and secondary rates subject to the sum-probability constraint
		\begin{maxi!}
			{\scriptstyle{\boldsymbol{r} \ge 0}}{\rho R_s + (1 - \rho) R_c}{\label{op:weighted_sum_rate}}{\label{ob:weighted_sum_rate}}
			\addConstraint{\sum_{i \in \mathcal{M}} r_i}{=1,}{\label{co:sum_probability}}
		\end{maxi!}
		where $\rho \in [0, 1]$ is the weight factor. By substituting \eqref{eq:secondary_rate} and \eqref{eq:primary_rate} into \eqref{ob:weighted_sum_rate}, the problem expands to
		\begin{maxi!}
			% {\scriptstyle{\boldsymbol{r} \ge 0, \boldsymbol{Q} \ge 0}}{R(\boldsymbol{r}, \boldsymbol{Q}) \triangleq \rho \sum_{i \in \mathcal{M}} r_i \log_2 \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr) \nonumber}{}{}
			% \breakObjective{\qquad \qquad \ + (1 - \rho) \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i}}
			{\scriptstyle{\boldsymbol{r} \ge 0, \boldsymbol{Q} \ge 0}}{\rho \sum_{i \in \mathcal{M}} r_i \log_2 \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr) \nonumber}{}{}
			\breakObjective{+ (1 - \rho) \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i}}{\label{ob:weighted_sum_rate_expanded}}
			\addConstraint{\sum_{i \in \mathcal{M}} r_i}{=1.}{}
		\end{maxi!}
		Similar to the Blahut-Arimoto algorithm \cite{Arimoto1972,Blahut1972}, we optimize $\boldsymbol{r}$ and $\boldsymbol{Q}$ alternatively to obtain the local optimal tag input distribution and the corresponding weighted sum-rate.
		\begin{theorem}
			For any valid tag input probability mass function $\boldsymbol{r}$ [$\boldsymbol{r} > \boldsymbol{0}$?], the optimal backward transition probability from channel output state $j$ to input state $i$ is given by
			\begin{equation}
				q_{j, i}^{\star} = \frac{r_i p_{i, j}}{\sum_{i'} r_{i'} p_{i', j}}.
				\label{eq:backward_transition}
			\end{equation}
			\label{th:backward_transition}
		\end{theorem}
		\begin{proof}
			Please refer to Appendix~\ref{ap:backward_transition}.
		\end{proof}
		\begin{theorem}
			For any valid backward transition probability $\boldsymbol{Q}$ [$\boldsymbol{Q} \ge \boldsymbol{0}$?], the optimal tag input probability mass function is given by
			\begin{equation}
				r_i = \frac{\bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr)^{\rho / (1 - \rho)} \prod_{j \in \mathcal{M}} q_{j, i}^{p_{i, j}}}{\sum_{i' \in \mathcal{M}} \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_{i'} \rvert^2 \bigr)^{\rho / (1 - \rho)} \prod_{j \in \mathcal{M}} q_{j, i'}^{p_{i', j}}}.
				\label{eq:input_distribution}
			\end{equation}
			\label{th:input_distribution}
		\end{theorem}
		\begin{proof}
			Please refer to Appendix~\ref{ap:input_distribution}.
		\end{proof}

	\end{section}

>>>>

			denotes the typical received signal energy over $N$ symbol blocks when the tag is at state $m$, which can be estimated at the user. For the ease of exposition, we sort $\{\sigma_m^2\}$ in a ascending order by bijective mapping $f \colon m \mapsto i$\footnote{Note that the mapping is not unique and the detection fails when two different constellation points yield the same energy level.} and let $\sigma_i^2 \le \sigma_j^2$ for $i < j$, $i,j \in \mathcal{M}$. Denote $\mathcal{H}_i$ as the hypothesis that the tag transmits $c_i$. Following \cite{Qian2019}, the optimal Maximum-Likelihood (ML) detector boils down to the energy detector, and the likelihood ratio between hypotheses $\mathcal{H}_i$ and $\mathcal{H}_j$ is

@@

				\Lambda_{i, j}(\boldsymbol{y}) \underset{H_j}{\overset{H_i}{\lessgtr}} 1 \iff z \underset{H_j}{\overset{H_i}{\lessgtr}} T_{i, j},

@@

			We then investigate the achievable rate of the backscatter link. Denote the tag input probability mass function as $\boldsymbol{r} = [r_1,\ldots,r_M]^T \in \mathbb{R}_+^{M \times 1}$ where $r_m$ is the probability to transmit symbol $c_m$ [strictly positive?]. For a given input distribution $\boldsymbol{r}$ and forward transition matrix $\boldsymbol{P}$, the achievable backscatter rate can be written as
			\begin{equation}
				R_c = \max_{\boldsymbol{Q} \ge 0} \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i} \ [\si{bpcu}],
				\label{eq:secondary_rate}
			\end{equation}
			where $\boldsymbol{Q} \in \mathbb{R}_{+}^{M \times M}$ is the backward transition probability matrix
			\begin{equation}
				\boldsymbol{Q} \triangleq
				\begin{bmatrix}
					q_{1, 1} & \ldots & q_{1, M} \\
					\vdots & \ddots & \vdots \\
					q_{M, 1} & \ldots & q_{M, M}
				\end{bmatrix},
			\end{equation}
			and $q_{j, i}$ denotes the transition probability from channel output state $j$ to input state $i$.
		\end{subsection}

@@

	\begin{section}{Tag Input Distribution}
		To achieve a flexible tradeoff between the backscatter and legacy links, we aims to optimize the tag input probability mass function, under $M$-PSK modulation, to maximize the weighted sum of the primary and secondary rates subject to the sum-probability constraint
		\begin{maxi!}
			{\scriptstyle{\boldsymbol{r} \ge 0}}{\rho R_s + (1 - \rho) R_c}{\label{op:weighted_sum_rate}}{\label{ob:weighted_sum_rate}}
			\addConstraint{\sum_{i \in \mathcal{M}} r_i}{=1,}{\label{co:sum_probability}}
		\end{maxi!}
		where $\rho \in [0, 1]$ is the weight factor. By substituting \eqref{eq:secondary_rate} and \eqref{eq:primary_rate} into \eqref{ob:weighted_sum_rate}, the problem expands to
		\begin{maxi!}
			% {\scriptstyle{\boldsymbol{r} \ge 0, \boldsymbol{Q} \ge 0}}{R(\boldsymbol{r}, \boldsymbol{Q}) \triangleq \rho \sum_{i \in \mathcal{M}} r_i \log_2 \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr) \nonumber}{}{}
			% \breakObjective{\qquad \qquad \ + (1 - \rho) \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i}}
			{\scriptstyle{\boldsymbol{r} \ge 0, \boldsymbol{Q} \ge 0}}{\rho \sum_{i \in \mathcal{M}} r_i \log_2 \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr) \nonumber}{}{}
			\breakObjective{+ (1 - \rho) \sum_{i \in \mathcal{M}} \sum_{j \in \mathcal{M}} r_i p_{i, j} \log_2 \frac{q_{j, i}}{r_i}}{\label{ob:weighted_sum_rate_expanded}}
			\addConstraint{\sum_{i \in \mathcal{M}} r_i}{=1.}{}
		\end{maxi!}
		Similar to the Blahut-Arimoto algorithm \cite{Arimoto1972,Blahut1972}, we optimize $\boldsymbol{r}$ and $\boldsymbol{Q}$ alternatively to obtain the local optimal tag input distribution and the corresponding weighted sum-rate.
		\begin{theorem}
			For any valid tag input probability mass function $\boldsymbol{r}$ [$\boldsymbol{r} > \boldsymbol{0}$?], the optimal backward transition probability from channel output state $j$ to input state $i$ follows Bayes' Theorem as
			\begin{equation}
				q_{j, i}^{\star} = \frac{r_i p_{i, j}}{\sum_{i'} r_{i'} p_{i', j}}.
				\label{eq:backward_transition}
			\end{equation}
			\label{th:backward_transition}
		\end{theorem}
		\begin{proof}
			Please refer to Appendix~\ref{ap:backward_transition}.
		\end{proof}
		\begin{theorem}
			For any valid backward transition probability $\boldsymbol{Q}$ [$\boldsymbol{Q} \ge 0$?], the optimal tag input probability mass function is given by
			\begin{equation}
				r_i = \frac{\bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_i \rvert^2 \bigr)^{\rho / (1 - \rho)} \prod_{j \in \mathcal{M}} q_{j, i}^{p_{i, j}}}{\sum_{i' \in \mathcal{M}} \bigl( 1 + \gamma \lvert h_{\mathrm{D}} + \sqrt{\alpha} h_{\mathrm{C}} c_{i'} \rvert^2 \bigr)^{\rho / (1 - \rho)} \prod_{j \in \mathcal{M}} q_{j, i'}^{p_{i', j}}}.
				\label{eq:input_distribution}
			\end{equation}
			\label{th:input_distribution}
		\end{theorem}
		\begin{proof}
			Please refer to Appendix~\ref{ap:input_distribution}.
		\end{proof}
		\begin{proposition}
			For any feasible initial point [constellation subset], the Algorithm is guaranteed to converge to the global optimal of problem \eqref{op:weighted_sum_rate} and achieve the maximum weighted-sum rate.
		\end{proposition}
		\begin{proof}
			The proof is similar to that in \cite[Chapter~9.3]{Yeung2008} and omitted here.
		\end{proof}
	\end{section}
